---
title: "Shipwrecks DB"
author: "Ben Best"
date: "December 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=T, eval=F)
```

##

See Other Bookmarks: 2016-09-30 shipwrecks lit db Michael Barrett

Google Oceans: Shipwrecks

- Ben would it be possible to add the Marine Protected Areas so we can estimate how many of the wrecks are in these areas. As you will see the database will have no more than 2000 records, but so far it is the first effort to digitalise them

- Iglika will add a column for the country names using the  ISO codes and an additional one using the United Nations Statistics Division codes. Next week hopefully Iglika and Lars will be able to run a quality control of the database since we have not done it yet. Please make any changes or comments to the manuscript version, if you feel we must expand on anything please let us know (we still need to work further on the Metadata).

- [Marine Ecoregions of the World](http://www.marineregions.org/maps.php?album=3264&pic=64936)
- [Country by EEZ](http://www.marineregions.org/maps.php?album=3264&pic=64931)
- distance from [shore](https://www.ngdc.noaa.gov/mgg/shorelines/gshhs.html)
- [depth](http://topex.ucsd.edu/WWW_html/srtm30_plus.html)
- [sea surface temperature](https://www.ncdc.noaa.gov/oisst)
- [chlorophyll](http://oceancolor.gsfc.nasa.gov/cms/data/viirs) ([data](http://oceandata.sci.gsfc.nasa.gov/VIIRS/Mapped/Cumulative/4km/CHL_chlor_a))
- [WDPA - World Database of Protected Areas - ProtectedPlanet.net](https://protectedplanet.net/)
- [Marine Regions - downloads](http://www.marineregions.org/downloads.php)
  - [Marine Regions - webservices](http://www.marineregions.org/webservices.php)
- [Large Marine Ecosystem - Digital Data](http://www.lme.noaa.gov/index.php?option=com_content&view=category&layout=blog&id=48&Itemid=268)

```{r libraries, eval=T}

# load libraries
suppressPackageStartupMessages({
  library(readxl)
  library(raster)
  library(rgeos)
  library(readr)
  library(tibble)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(mregions) # devtools::install_github("ropenscilabs/mregions")
  library(leaflet)
  library(htmltools)
  library(maps)
  library(mregions)
  library(ggplot2)
  library(sp)
  library(sf)
  library(rgdal)
  library(ncdf4)
  library(scales)
  library(robis)
  library(DT)
})
select = dplyr::select

# projections
crs_mol = '+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs'
crs_gcs = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
```


```{r points, eval=T}
# spatial points
wrecks_df = read_excel('data/shipwrecks_db_1.xlsx', na='NA')
# rename so not having duplicate of "ISO Country Code" # paste(names(wrecks), collapse="','")
names(wrecks_df) = c('num','name','year','type','country','subcountry','water_body','depth_str','lat','lon',
  'accuracy','last_updated','source','notes','iso','iso_num','iso_sub')

# create points, filter NAs
wrecks_df = wrecks_df %>%
  filter(!is.na(lon), !is.na(lat)) # 3 NA's for lon
wrecks_pts = wrecks_df
coordinates(wrecks_pts) = ~lon+lat
proj4string(wrecks_pts) = CRS(crs_gcs)
st_write(st_as_sf(wrecks_pts), './results/shipwrecks_pts.shp')

# table
wrecks_df %>%
  mutate(
    source_link = sprintf('<a href="%s">source</a>', source)) %>%
  select(-notes, -source) %>%
  datatable(escape=F)
```


```{r map, eval=T}
leaflet(wrecks_pts) %>%
  addProviderTiles('Esri.OceanBasemap') %>%  # or 'Esri.OceanBasemap'/'Stamen.TonerLite'
  # see [all providers](http://leaflet-extras.github.io/leaflet-providers/preview/index.html)
  addMarkers(
    clusterOptions = T,
    popup = ~sprintf("<strong>%s</strong> (%d)<br>type: %s<br>num: %d", name, year, type, num))
```


```{r land}
depth_nc    = '../obis-lat-time-fig/data/GEBCO_2014_2D.nc'

depth_r = raster(depth_nc, layer = 'elevation')

system.time({
  wrecks_elev = raster::extract(depth_r, wrecks_pts, method='bilinear', df=T)}) # 48.090 seconds
# Elevation.relative.to.sea.level

hist(wrecks_pts$Elevation.relative.to.sea.level)
summary(wrecks_pts$Elevation.relative.to.sea.level)
```


```{r marine regions, eval=F}
# This seems to time out and has complicated results when run on filtered subset

geocode = function(df){ mregions::mr_rev_geo_code(df$lat, df$lon) }

wrecks_mregions_csv = './results/shipwrecks_mregions.csv'

if (!file.exists(wrecks_mregions_csv)){
  wrecks_df %>%
    #filter(num %in% 29:32) %>% # testing
    group_by(num, name) %>%
    nest() %>%
    mutate(
      mregions = purrr::map(data, geocode)) %>%
    select(-data) %>%
    unnest() %>%
    write_csv(wrecks_mregions_csv)
  # looks unique by MRGID: gazetteerSource, placeType, preferredGazetteerName
}

# read data
wrecks_mregions = read_csv(wrecks_mregions_csv)

# summarize by wrecks
wrecks_mregions %>%
  group_by(num, name) %>%
  summarize(
    n_gazetteerSource = n_distinct(gazetteerSource),
    n_placeType       = n_distinct(placeType)) %>%
  datatable()

# summarize by gazetteerSource
wrecks_mregions %>%
  group_by(gazetteerSource) %>%
  summarize(
    n_wrecks = n_distinct(num)) %>%
  datatable()

# summarize by gazetteerSource, placeType
wrecks_mregions %>%
  group_by(gazetteerSource, placeType) %>%
  summarize(
    n_wrecks = n_distinct(num)) %>%
  datatable()
```

## Shapefiles

Using PostGIS since big features.

Installed:

- [Postgres.app](http://postgresapp.com/) – the easiest way to get started with PostgreSQL on the Mac

- [pgAdmin](https://www.postgresql.org/ftp/pgadmin3/pgadmin4/v1.1/macos/) GUI

```bash
sudo chown -R `whoami`:admin /usr/local/share
brew link lzlib
brew install postgis
# start manually
pg_ctl -D /usr/local/var/postgres -l /usr/local/var/postgres/server.log start
# stop manually
pg_ctl -D /usr/local/var/postgres stop -s -m fast
# connect to default postgres database
psql postgres
```

Enable PostGIS

```psql
CREATE DATABASE shipwrecks;
-- DROP DATABASE shipwrecks;
\connect shipwrecks
-- Enable PostGIS (includes raster)
CREATE EXTENSION postgis;
-- Enable Topology
CREATE EXTENSION postgis_topology;
-- Enable PostGIS Advanced 3D
-- and other geoprocessing algorithms
-- sfcgal not available with all distributions
CREATE EXTENSION postgis_sfcgal;
-- fuzzy matching needed for Tiger
-- CREATE EXTENSION fuzzystrmatch;
-- rule based standardizer
-- CREATE EXTENSION address_standardizer;
-- example rule data set
-- CREATE EXTENSION address_standardizer_data_us;
-- Enable US Tiger Geocoder
-- CREATE EXTENSION postgis_tiger_geocoder;
```

Per [edzer/sfr: Simple features for R](https://github.com/edzer/sfr), installed dev version to get `st_write()` compatability with PostGIS vs:

```
st_write(ply, sprintf('PG:dbname=%s', db), obj, layer_options = 'OVERWRITE=true')
Error in guess_driver(dsn) :
  guess_driver: PostgreSQL not available in supported drivers, see `st_drivers()'
```

uninstalling:

sudo rm -rf /Library/Frameworks/Python.framework/

- [osx - What is the safest way to removing Python framework files that are located in different place than Brew installs - Stack Overflow](http://stackoverflow.com/questions/22774529/what-is-the-safest-way-to-removing-python-framework-files-that-are-located-in-di)


So installing:

- [GDAL 2 On Mac with Homebrew](http://www.karambelkar.info/2016/10/gdal-2-on-mac-with-homebrew/)

```bash
# unlink gdal 1.x
brew unlink gdal
brew prune gdal

# remove kyngchaos.com GDAL framework install
sudo rm -rf /Library/Frameworks/GDAL.framework/
brew prune

# tap into osgeo4mac
brew tap osgeo/osgeo4mac && brew tap --repair

# install gdal2
#brew install gdal2 --with-armadillo \
#  --with-complete --with-libkml --with-unsupported
# Error: undefined method `rebuild' for #<BottleSpecification:0x007fa31d8a1bd8>
brew install gdal2

# link gdal2
brew link --force gdal2

# verify version in new terminal
gdal-config --version
# 2.1.1
gdal-config --libs
# -L/usr/local/Cellar/gdal2/2.1.1/lib -lgdal
gdal-config --cflags
# -I/usr/local/Cellar/gdal2/2.1.1/include
```

- reinstall rgdal

```R
remove.packages('rgdal')
install.packages('rgdal')
library(rgdal) # test
remove.packages('rgeos')
install.packages('rgeos')
```

- restart R session

```bash
brew info gdal2
```

```R
library(devtools)
Sys.setenv(GDAL_CONFIG='/usr/local/opt/gdal2/bin/gdal-config')
install_github('edzer/sfr')
```

screw it:

```R
install.packages("sf")
```

# load shp into db

[shp2pgsql: Using the ESRI Shapefile Loader](http://postgis.net/docs/using_postgis_dbmanagement.html#shp2pgsql_usage)

```bash
ogrinfo
cd /Users/bbest/github/shipwrecks
shp2pgsql -i -D -d -s 4269 -I ./data/shipwrecks_pts.shp wrecks | psql -d shipwrecks
shp2pgsql -i -D -d -s 4269 -I ./data/World_EEZ_v9_20161021/eez.shp eez | psql -d shipwrecks
shp2pgsql -i -D -d -s 4269 -I ./data/WDPA_Dec2016-shapefile-polygons.shp wdpa | psql -d shipwrecks
shp2pgsql -i -D -d -s 4269 -I ./data/MEOW/meow_ecos.shp meow | psql -d shipwrecks
shp2pgsql -i -D -d -s 4269 -I ./data/LME66/LME66.shp lme | psql -d shipwrecks
```

```bash
brew install qscintilla2
brew install qgis2
```


```{r shapefiles}
library(sf)
library(RPostgreSQL)

shapes = list(
  wrecks = './results/shipwrecks_pts.shp',
  wdpa   = './data/WDPA_Dec2016-shapefile/WDPA_Dec2016-shapefile-polygons.shp', # 213,328 features
  eez    = './data/World_EEZ_v9_20161021/eez.shp',
  lme    = './data/LME66/LME66.shp',
  iho    = './data/IHO/World_Seas.shp')

shp = shapes[[2]]
obj = names(shapes[2])

db = 'shipwrecks'

con = dbConnect(dbDriver("PostgreSQL"), dbname=db, host='localhost')


ply = st_read(shp)
st_layers("PG:dbname=shipwrecks")
st_read("PG:dbname=shipwrecks")
eez = st_read_db(con, 'eez')
plot(eez)
st_write(ply, sprintf('PG:dbname=%s', db), obj, layer_options = 'OVERWRITE=true')


#select from wrecks, eez where st_intersects(wrecks.geom, eez.geom) and num in (29,30,31);

# test if column is unique
#select mrgid from eez group by mrgid having count(*) > 1;
#select wdpaid from wdpa group by wdpaid having count(*) > 1;


system.time({
  wrecks_wdpa = dbGetQuery(
    con, 
    "SELECT 
      wrecks.num AS wreck_num, wrecks.name AS wreck_name, 
      wdpa.gid AS wdpa_gid, wdpa.name AS wdpa_name 
    FROM wrecks, wdpa WHERE ST_INTERSECTS(wrecks.geom, wdpa.geom);") %>%
    dbGetQuery("SELECT * FROM wdpa")
    arrange(wreck_num) %>%
    write_csv('./results/shipwrecks_wdpa.csv')
})
#  user  system elapsed 
# 0.010   0.002   4.205
#read_csv('./data/shipwrecks_wdpa.csv') %>% head
# multiple WDPAs per single wreck (eg wreck_num=23)
#read_csv('./data/shipwrecks_wdpa.csv') %>% dim # 1879

# write postgres table to csv for joining
tbl(src_postgres(dbname=db, host='localhost'), 'wdpa') %>%
  select(-geom, -name) %>%
  rename(wdpa_gid = gid) %>%
  as_data_frame() %>%
  write_csv('./data/wdpa.csv')

# left join wdpa info to intersection
read_csv('./results/shipwrecks_wdpa.csv') %>%
  left_join(
    read_csv('./data/wdpa.csv'),
    by='wdpa_gid') %>%
  write_csv('./results/shipwrecks_wdpa2.csv')

# write postgres table to csv for joining
tbl(src_postgres(dbname=db, host='localhost'), 'wdpa') %>%
  select(-geom, -name) %>%
  rename(wdpa_gid = gid) %>%
  as_data_frame() %>%
  write_csv('./data/wdpa.csv')

# left join wdpa info to intersection
read_csv('./results/shipwrecks_wdpa.csv') %>%
  left_join(
    read_csv('./data/wdpa.csv'),
    by='wdpa_gid') %>%
  write_csv('./results/shipwrecks_wdpa2.csv')

system.time({
  dbGetQuery(
    con, 
    "SELECT 
      num AS wreck_num, name AS wreck_name, 
      mrgid AS eez_mrgid, geoname AS eez_geoname 
    FROM wrecks, eez WHERE ST_INTERSECTS(wrecks.geom, eez.geom);") %>%
    arrange(wreck_num) %>%
    write_csv('./data/shipwrecks_eez.csv')
})
#  user  system elapsed 
# 0.018   0.002   7.547
# read_csv('./data/shipwrecks_eez.csv') %>% head

# vs SLOW...
# eez    = st_read(shapes[['eez']])
# wrecks = st_read(shapes[['wrecks']])
# 
# system.time({
#   wrecks %>%
#     st_intersects(eez) %>%
#     select(num, geoname)
# })


ply = rgdal::readOGR(dirname(shp), tools::file_path_sans_ext(basename(shp)))

# for smaller data, read in shapefile
wrecks_pts = st_read('./results/shipwrecks_pts.shp')
lme_ply    = st_read('./data/LME66/LME66.shp')
meow_ply   = st_read('./data/MEOW/meow_ecos.shp')

w = wrecks_pts %>%
  st_intersects(lme_ply, sparse=F) %>% # return sparse logical matrix with rows and columns corresponding to features x and y
  as_tibble()

w2 = wrecks_pts %>%
  st_intersection(lme_ply) %>% # return sparse logical matrix with rows and columns corresponding to features x and y
  as_tibble()

leaflet::dra
st_intersects2() = function(st_intersects(...))

w %>%
  rownames_to_column('wreck_row') %>%
  gather(lme_name, intersects, -wrecks_row) %>%
  mutate(
    wrecks_row = as.integer(wrecks_row),
    lme_row = as.integer(stringr::str_replace(lme_name, 'V',''))) %>%
  filter(intersects==T) %>%
  left_join(
    wrecks_pts %>% as('data.frame') %>%
      mutate(
        wrecks_pts %>% 
          as_tibble() %>%
          select(-geometry) %>%
          mutate(
            wreck_row = row_number())
      )
  )
  tail
  
library(rgeos)

gIntersects()
intersect()     
         
w_l %>%
  head


w_l %>% flatten_df()
  
w_l %>%
  head
  select(num, geoname)


pts_ply = sp::over(
  st_read()
  wrecks_pts[28:31,], ply)
pts_ply

```

```{r raster}
# raster::extract()
dir_asc = './data/BioOracle'

rasters = stack(list.files(dir_asc, '\\.asc$', full.names=T))
crs(rasters) = crs_gcs

wrecks = st_read('./results/shipwrecks_pts.shp') %>% as('Spatial')

system.time({
  raster::extract(rasters, wrecks, df=T) %>%
    write_csv('./results/shipwrecks_rasters.csv')
})
 #   user  system elapsed 
 # 27.159   2.399  29.737 

```
